---
title       : Data and Models
subtitle    : Data Visualization, Week 04
author      : Kieran Healy
job         : Duke University
framework   : revealjs        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : solarized_light      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
url         : {lib: "."}
revealjs:
  theme: night
  transition: fade
  center: "true"
  height: 600
---

<style type='text/css'>
.reveal {
  font-size: 28px;  
}
</style>

```{r Setup, include=FALSE, results='hide', warning=FALSE}

library(devtools)
library(ggplot2)

opts_chunk$set(fig.path="assets/fig/wk04-fig-",
               cache.path="../cache/wk-04-working-",
               dev=c("png","pdf"),
               fig.width=6,
               fig.height=5,
               dpi=300,
               fig.show="hold",
               fig.lp="fig:",
               cache=FALSE,
               par=TRUE,
               echo=TRUE,
               message=FALSE,
               warning=FALSE)


```


# Data and Models

## Data Visualization, Week 4

### Kieran Healy, Duke University

---

## Outline for Today

0. Housekeeping
1. Pairs Plots
2. Looking at Data while Modeling
3. `broom`
4. A (curse) word about Git
5. The NZ Electorate data

--- 

## How to Navigate these Slides

- When you view them online, notice the compass in the bottom right corner
- You can go left or right, or sometimes down to more detail.
- Hit the `Escape` key to get an overview of all the slides. On a phone
  or tablet, pinch to get the slide overview.
- You can use the arrow keys (or swipe up and down) in this view, as well. 
- Hit `Escape` again to return to the slide you were looking at. 
- On a phone or tablet, tap the slide you want.

---


## Organs Data

```{r, get-organdata, echo= TRUE}

organs.url <- "https://raw.githubusercontent.com/socviz/soc880/gh-pages/data/organ-procurement.csv"
organs.data <- read.csv(url(organs.url), row.names=1)
dim(organs.data)
head(organs.data)

## For convenience, let R know year is a time measure.
organs.data$year <- as.Date(strptime(organs.data$year, format="%Y"))

```

---

#### Pairs Plot

```{r, warning=FALSE}

## install.packages("GGally")
library(GGally)

my.vars <- c("donors", "pop", "gdp",
             "pubhealth", "assault")

#pdf(file="../figures/ggpairs-plot-1.pdf", height=15, width=15)
p <- ggpairs(subset(organs.data, select=my.vars),
        upper=list(continuous="smooth"),
        lower=list(continuous="points"))
print(p)
#dev.off()

```

---

#### Generalized Pairs Plots 

```{r, warning=FALSE}


my.vars <- c("donors", "pop.dens", "gdp",
             "pubhealth", "world", "consent.law",
             "assault", "opt")

# pdf(file="../figures/ggpairs-plot-2.pdf", height=20, width=20)
p <- ggpairs(subset(organs.data, select=my.vars),
        upper=list(continuous="smooth", combo="box"),
        lower=list(continuous="points", combo="dot"))
print(p)
# dev.off()

```

---

## Looking While Modeling

- Frank Harrell, *Regression Modeling Strategies* (Springer 2001) is an older text, but the visualization principles are good. Two of his useful software libraries for R are `Hmisc` and `rms`.

```{r}

### install.packages("Hmisc")
### install.packages("rms")

library(Hmisc)
library(rms)

vc <- varclus(~ donors + pop.dens + gdp + pubhealth + assault + world + consent.law,
              data=organs.data)
plot(vc)

```

---

```{r}

plot(naclus(organs.data))

```

---

```{r}

s <- summary(donors ~ pop.dens + gdp + pubhealth + assault + world + consent.law,
              data=organs.data)

plot(s, cex.labels = 0.6)

```

---

## Diagnostic Plots for a Poorly-Specified Model

```{r}

out.lm <- lm(donors ~ pop.dens + gdp + pubhealth + assault + consent.law,
             data=organs.data)

summary(out.lm)

```

---

```{r}

plot(out.lm, which=1)

```

---

```{r}

plot(out.lm, which=2)

```


---

```{r}

plot(out.lm, which=3)

```

---

```{r}

plot(out.lm, which=5)

```

---

### Plotting results with Hmisc

```{r}

## NB
out.ols <- ols(donors ~ pop.dens + gdp + pubhealth + assault + consent.law,
             data=organs.data)

d <- datadist(organs.data)
options(datadist="d")
options(digits = 2)

plot(summary(out.ols))

```

---

## Tidy Model Output with `broom`

```{r}

## install.packages("broom")

library(broom)

## more irresponsible model
out.lm <- lm(pubhealth ~ pop.dens + gdp + country,
             data=organs.data)

summary(out.lm)


```

---

### Model objects have a lot of information in them

```{r}

str(out.lm)

```

---

### `broom` turns summary model output into data frames

```{r}

out.tidy <- tidy(out.lm, conf.int = TRUE)
class(out.tidy)

head(out.tidy)

```

---

### It will also "augment" data frames

```{r}

out.aug <- augment(out.lm)
head(out.aug)


```

---

### Tidy output is easy to quickly plot

```{r}

p <- ggplot(out.tidy, aes(y=estimate,
                          x=reorder(term, estimate)))
p + geom_pointrange(aes(ymin=conf.low,
                        ymax=conf.high)) +
    geom_hline() +
    coord_flip()

```

---

### Tidiers are available for many packages

```{r}

library(survival)

out.surv <- survfit(coxph(Surv(time, status) ~ age + sex, data=lung))

out.tidy <- tidy(out.surv)

p <- ggplot(out.tidy, aes(time, estimate))

p + geom_line() +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = .2)

```

---

### Model objects have associated `summary` and `plot` methods

---

### New Zealand Electorate Data

- On GitHub: https://github.com/kjhealy/nz-sex-ratios

---

## Excursus on Git

---

### Git is Horrible

- But it can help you keep track of what you did
- It's also where a lot of software development happens
- So it's worth getting a feel for it

---

### Simplest GitHub Use Case

- You have no account there.
- You download stuff from it.

---

### Next Simplest GitHub Use Case

- You have an account there.
- You fork (i.e., copy) things on other people's GitHub accounts to
  your GitHub account. 
- Then you clone (i.e., copy, or download) them to your laptop.
- This kind of confusing, doubled-up, garbage terminology is typical of Git. 

--- 

### Basic One-Person Regular Git Use Case

- You have a file under version control.
- You make a change to it and save the file.
- You stage or `add` a snapshot of the changed files, 
- Then you `commit` the snapshot of changes to the
repository, with a message saying what you did.

---


### Basic One-Person Regular Git Use Case

- You have a file under version control.
- You make a change to it and save the file.
- You stage or `add` a snapshot of the changed files, 
- Then you `commit` the snapshot of changes to the
  repository, with a message saying what you did.
- You `push` that change up to e.g. GitHub.

---

### Tutorial

- https://try.github.io/levels/1/challenges/1

---

## Back to the New Zealand Data

---

